---
title: "Paper 2 - Design Intervention"
author: "Lee Doucet"
date: "16/02/2021"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intervention Design 

	The Ontario Provincial government is undertaking an initiative to reopen 
select restaurants for in-person dining to determine the relationship between 
in-person dining and the spread of Covid-19 in the community. As Ontario has 
moved through the peak of the second wave and total daily case counts of 
Covid-19 are on the decline, Ontario Health believes this is a reasonably safe 
opportunity to conduct an intervention that tests the impact of reopening
in-person dining on future Covid-19 cases in the community. The intervention 
will take place as a Random Control Trial (RCT) that divides all restaurants 
in Ontario into two groups by a simple random sample without replacement 
(SRSWOR).

	Health officials will conduct this intervention for two-weeks from March 1st 
until March 15th. Once this time period has elapsed, all in-person dining will
be closed across the province while health officials monitor for any instances 
of community spread of Covid-19 where restaurants were open for in-person dining 
for two-weeks. This time frame was chosen as it is consistent with the current
medical knowledge of anywhere from 5 to 14 days from the initial date of 
infection to the onset of symptoms of Covid-19 to appear in individuals who are 
not asymptomatic. Data from this intervention will inform both Ontario Health 
and Public Health Ontario on next courses of action and whether to expand the 
intervention or not.

```{r}
```

## Reason for Random Control Trials  
 
	The intervention will use the RCT method as it is the most appropriate in the
context of the experiment. All restaurants in Ontario will divided into two 
groups. One group will be the treatment group which will be allowed to reopen 
their in-person dining and the other group will be a control group which will 
have no change and their in-person dining will remain closed. It’s imperative 
that no bias be present when selecting who is eligible for treatment as there 
are financial benefits to the restaurants that are sorted into the treatment 
group. The RCT will give each restaurant the same probability of being chosen 
randomly to be selected for the treatment or being placed in the control group. 
This methodology is both statistically fair while providing a transparent and 
simple enough process that the general public can understand the reasoning 
behind why only certain restaurants are being allowed to open for in-person 
dining.The usefulness of RCT’s have made them the “gold standard of impact 
evaluation” (Gertler et al, 64). 

	An important element to a robust RCT is having a large number of units to 
ensure the treatment group and control group are statically identical. This 
does not mean that each group needs to be prefect clones of each other 
(Alexander, Hunt Data), as this would make statistical analysis impossible in 
this case as restaurants are not identical. They vary in size, location, and 
serve different types of communities. There are also characteristics that are 
more challenging to observe such as personal motivations and willingness to 
follow proper protocols with Covid-19 compliance. What is important is the 
average of those properties are distributed as equal as possible between the 
treatment and control groups. The larger the size of the groups that were 
randomly sorted, the higher probability this will be the case. This is a benefit
of selecting an RCT for the intervention, with the population being all legal 
restaurants in Ontario, there is a sufficient high probability of both groups 
being statically identical.


